---
title: "Read Data 101 - include only primary decision makers"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

Data from here:
https://osf.io/8kzfv?view_only=d05ee82e58dc4c0f84448785ab839bb3
Original methods and full paper here:
https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0265662

# Load required packages
```{r loadpkgs, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE, results='hide'}
figures <- "figures/"
library(stringr)
library(dplyr)
library(DescTools)
library(car)
require(foreign)
require(ggplot2)
require(Hmisc)
require(reshape2)
library(tidyverse)
library(broom)
library(aod)
library(glmtoolbox)
library(fmsb)
library(uwot)
library(ggthemes)
library(ggforce)
library(readxl) #leaves characters as characters
library(ROCR) # Use ROCR package to plot ROC curve & AUC
library(pROC)
library(caret)
library(xgboost)
library(smotefamily) # one method of ML data augmentation
library(rsample)
library(mltools) # Wrapper allows UMAP calculations
library(data.table)
library(heatmaply)
library(rstatix) # to calculate Kendall correlation with associated p-values.
library(IHW) # Bioconductor package to calculate BH threshold
library(dbscan)
library(corrplot)
library(showtext)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
library(splines)
library(cluster)
library(factoextra)
library(clusterSim)
library(RColorBrewer)
```

# Open Excel file
```{r read_data_1, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
#Save the data so can run subsequent analysis ad lib
SAVE=TRUE

#quicker version readxl
ml <- read_excel('Canine health results.xlsx', sheet = "All")
cols <- c("Col_a")
for (col in 2:ncol(ml)){
  if (col <= 26){
    cols <- cbind(cols,paste0("Col_",letters[col]))
  } else if (col <= 52) {
    cols <- cbind(cols,paste0("Col_a",letters[col-26]))
  } else {
    cols <- cbind(cols,paste0("Col_b",letters[col-52]))
  }
}
names(ml) <- cols

```

Rename variables
```{r RENAME_MATCH_2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- dplyr::rename(ml, unique.resp.number = Col_a)
ml <- dplyr::rename(ml, Location = Col_c)
ml <- dplyr::rename(ml, setting = Col_e)
ml <- dplyr::rename(ml, Education = Col_g)
ml <- dplyr::rename(ml, Animal_Career = Col_h)
ml <- dplyr::rename(ml, Income = Col_i)
ml <- dplyr::rename(ml, C_Age = Col_j)
ml <- dplyr::rename(ml, C_Gender = Col_k)
ml <- dplyr::rename(ml, C_Diet = Col_l)
ml <- dplyr::rename(ml, Size = Col_o)
ml <- dplyr::rename(ml, D_Age = Col_p)
ml <- dplyr::rename(ml, D_Gender = Col_r)
ml <- dplyr::rename(ml, Therapeutic_Food = Col_s)
ml <- dplyr::rename(ml, Visits = Col_t)
ml <- dplyr::rename(ml, Vet_Health = Col_u)
ml <- dplyr::rename(ml, Meds = Col_ag)
ml <- dplyr::rename(ml, Health = Col_ah)
ml <- dplyr::rename(ml, D_Diet = Col_ai)
ml <- dplyr::rename(ml, Decision_Maker = Col_at)
head(ml)
```

# Start by creating separate Sex and neuter status variables
```{r split_gender_Neuter, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Sex <- rep("Male",nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('Male', ml$D_Gender[i], fixed = TRUE)){
       ml$D_Sex[i] <- "Male"
       } else{
       ml$D_Sex[i] <- "Female"
       }
}

#Checks and Balances
dplyr::count(ml, D_Sex)

ml$D_Neuter <- rep("BLANK",nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('intact', ml$D_Gender[i], fixed = TRUE)){
       ml$D_Neuter[i] <- "Intact"
       }
    else {
       ml$D_Neuter[i] <- "Neutered"
    }
}
dplyr::count(ml, D_Neuter)
```

# check D_Age
```{r check_D_Age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, D_Age)
```

# Dog Age remove <1 years and unsure, then change to numeric
```{r tidy_D_Age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <-subset(ml, D_Age != "<1" & D_Age != "Unsure")
ml$D_Age<-as.numeric(ml$D_Age)
dplyr::count(ml, D_Age)
```

# Initial Visualisation Excluding Health indicators with UMAP
```{r PrepForUMAP, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE,fig.showtext=TRUE}
uml<-ml

#Temporarily create the Owner Diet data labels for use later.

#Numeric -Euclidean
D_Age <- uml$D_Age
C_Age <- factor(uml$C_Age, 
                ordered=TRUE,levels=c("18-19",
                "20–29", "30–39", "40–49", 
                "50–59", "60–69",">70"))

#Ordered FALSE since TOY is not strictly smaller than Small, but different.
Size <- factor(uml$Size, ordered = FALSE, levels=c(
        "Toy (≤ 4.1 kg or ≤ 9 lb)",
"Small (4.5–10.9 kg or 10–24 lb)",
"Medium (11.4–22.3 kg or 25–49 lb)",
"Large (22.7–40.5 kg or 50–89 lb)",
 "Giant (> 40.9 kg or > 90 lb)") )

uml$Education <- factor(uml$Education, 
                   ordered=TRUE,
                   levels=c("Did not complete high school" ,
                      "High school or equivalent",
                      "College or University award lower than undergraduate degree",
                      "University undergraduate degree",
                      "University postgraduate degree lower than doctoral degree e.g. masters",
                      "Doctoral degree"))
                      
C_DietLabs<-uml$C_Diet
D_DietLabs<-uml$D_Diet
Edu_Labs <- uml$Education

Education <- uml$Education
#Not truely ordered because some "prefer not to say"
Income <- as.factor(uml$Income)

#Factors even if binary - use Hamming
D_Sex <- as.factor(uml$D_Sex)
C_Gender <- as.factor(uml$C_Gender)
Neuter_Status <- as.factor(uml$D_Neuter)
Decision_Maker <- as.factor(uml$Decision_Maker)
setting <- as.factor(uml$setting)
Animal_Career <- as.factor(uml$Animal_Career)
Location <- as.factor(uml$Location)  
D_Diet <- as.factor(uml$D_Diet) 
C_Diet <- as.factor(uml$C_Diet) 

#Reforming so we can use the explicitly set data types...
#Numeric Euclidean
#Ordered Manhattan
#Factor Hamming
#Factor Hamming
#Factor Hamming
Udata <- data.frame(
  cbind(D_Age, 
        C_Age, Education,
        Size,  
        Income, Neuter_Status,  D_Sex, C_Gender, 
        Animal_Career, C_Diet, D_Diet,
        setting, Decision_Maker, Location),
  stringsAsFactors = TRUE) #Should be no strings as already factors
#Labels

#if using uwot
targets <- c(rep("euclidean",1),
             rep("manhattan", 2),
             rep("hamming", 11) )

Udata$C_DietLabs <- C_DietLabs
uKeeper <- complete.cases(Udata)
Udata <- Udata[uKeeper,]

CLabs <- Udata$C_DietLabs
Edu_Labs <- Edu_Labs[uKeeper]
#For other applications:
write.csv(Udata,"Udata.csv")
# Dont include the labels themselves in the UMAP
Udata <- dplyr::select(Udata,-C_DietLabs)

# nice-ify labels
CLabs <- stringr::str_replace(CLabs,"Vegetarian \\(consuming plants, eggs and milk\\)","Vegetarian")
CLabs <- stringr::str_replace(CLabs,"Omnivore \\(i.e. eat meat\\)","Omnivore")
CLabs <- stringr::str_replace(CLabs,"Omnivore reducing animal product consumption","Omnivore \\(Reducing\\)")
CLabs <- stringr::str_replace(CLabs,"Vegan \\(consuming no animal products\\)" ,"Vegan")
CLabs <- stringr::str_replace(CLabs,"Pescatarian \\(including fish but no other meats\\)","Pescatarian")

```
```{r}
#Set up for doggie   
Labs <- D_DietLabs[uKeeper] #remember we complete cased above!

# nice-ify labels
Labs <- stringr::str_replace(Labs,"Meat-based – conventional","Meat-based")
Labs <- stringr::str_replace(Labs,"Vegan \\(consuming no animal products\\)","Vegan")
Labs <- stringr::str_replace(Labs,"Meat-based – lab-grown","Lab-grown")
Labs <- stringr::str_replace(Labs,"Meat-based – raw","Raw food")
Labs <- stringr::str_replace(Labs,"Vegetarian \\(including eggs or milk, but not meat\\)","Vegetarian")

```

#Perform UMAP run
```{r drawoneumap}
# Use UWOT library rather than UMAP since has built in categorical support
# hyperparameters, largely defaults

  n_epochs = 50000
  n_components = 2
  n_n = 20
  learning_rate = .01
  scale = "maxabs"
  spread = 1
  min_dist = 0.001
  set_op_mix_ratio = 1
  local_con = 1
  bandwidth = 1
  repulsion_strength = 1
  neg = 5
  n_trees = 50
  target_weight = 0.5 
  init="normlaplacian"

  set.seed(666)
  um <- uwot::umap(Udata, target_metric=targets,
                   spread=spread, n_neighbors = n_n,
                   min_dist = min_dist, negative_sample_rate = neg,
                   n_epochs = n_epochs, local_connectivity = local_con, 
                   n_trees=n_trees, scale=scale
  )   
 df=as.data.frame(cbind(um[,1],um[,2]))
  names(df) <- c("x","y") 
 ggplot(df, aes(x, y)) +
    geom_point(aes(fill = Labs),alpha=0.8,shape=21,size=2, stroke=0.2) + labs(fill="Dog Diet")+
    xlab("UMAP X co-ordinates")+ylab("UMAP Y co-ordinates")+
    #geom_segment(aes(y=bottom,yend=top,x=-Inf,xend=-Inf))+
    #geom_segment(aes(y=-Inf,yend=-Inf,x=left,xend=right))+
    #scale_y_continuous(breaks=seq(bottom,top,2))+
    #scale_x_continuous(breaks=seq(left,r-ight,2))+
    geom_rangeframe() + theme_tufte(base_family = "Gill Sans") 
    fname=paste0("umaps/","i", i,"min_dist",min_dist,
               "_spread" ,spread,"_nn" ,n_n, "_lcon" ,
               local_con,"_neg",neg,"_epochs",n_epochs, "_trees", n_trees,"_", scale,".png")
    #Test Umap plot including hyperparameters
    #ggsave(fname)
```
#Perform density clustering analysis
##And Dog clustering metrics
```{r clustering, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
suppressWarnings(file.remove("dog_diet_clusters.csv"))
set.seed(666)
db <- dbscan(um, eps=0.7, minPts = 50) #eps=1 minPts=5 db <- dbscan(um, eps=0.8, minPts = 25) dbscan(um, eps=0.86, minPts = 46)
print(db)
combi <- cbind(df,db$cluster,Labs)
names(combi) <- c("x","y","cluster","Diet")
print(paste("total observations = ", nrow(combi) ) )
output <- dplyr::count(combi, Diet, name="All Data")
print(output)
write.table(output,
            "dog_diet_clusters.csv",
            sep=",",
            row.names=FALSE,
            append = TRUE)
centeroids <- data.frame(mx=as.numeric(), my=as.numeric())
for (CLUSTER in seq(0, max(db$cluster)) ){
  print(paste("cluster",CLUSTER) )
  this_clust <- subset(combi,cluster==CLUSTER)
  print(paste("total observations in cluster", CLUSTER," = ", nrow(this_clust) ))
  output <- dplyr::count(this_clust, Diet, name=paste("Cluster", CLUSTER))
  print(output)
  write.table(output,
            "dog_diet_clusters.csv",
            sep=",",
            row.names=FALSE,
            append = TRUE)

  if (CLUSTER >0){
    mx <- mean(this_clust$x)
    my <- mean(this_clust$y)
    lc <- data.frame(mx=mx,my=my)
    centeroids <- rbind(centeroids,lc)
    print(paste("Cluster", 
                CLUSTER, 
                "centeroid is",
                round(mx,2),round(my,2)))
  }
}
```

##Owner Diet clustering metrics
```{r ownercluster data}
suppressWarnings(file.remove("owner_diet_clusters.csv"))
combi <- cbind(df,db$cluster,CLabs)
names(combi) <- c("x","y","cluster","C_Diet")
print(paste("total observations = ", nrow(combi) ) )
output <- dplyr::count(combi, C_Diet, name="All Data")

print(output)
write.table(output,
            "owner_diet_clusters.csv",
            sep=",",
            row.names=FALSE,
            append = TRUE)

for (CLUSTER in seq(0, max(db$cluster)) ){
  print(paste("cluster",CLUSTER) )
  this_clust <- subset(combi,cluster==CLUSTER)
  print(paste("total observations in cluster", CLUSTER," = ", nrow(this_clust) ))
  output <- dplyr::count(this_clust, C_Diet, name=paste("Cluster", CLUSTER))
  print(output)
  write.table(output,
            "owner_diet_clusters.csv",
            sep=",",
            row.names=TRUE,
            append = TRUE)
}


```


##Owner Education clustering metrics
```{r Educluster data}
suppressWarnings(file.remove("owner_edu_clusters.csv"))
combi <- cbind(df,db$cluster,Edu_Labs)
names(combi) <- c("x","y","cluster","Education")
print(paste("total observations = ", nrow(combi) ) )
output <- dplyr::count(combi, Education, name="All Data")

print(output)
write.table(output,
            "owner_edu_clusters.csv",
            sep=",",
            row.names=FALSE,
            append = TRUE)

for (CLUSTER in seq(0, max(db$cluster)) ){
  print(paste("cluster",CLUSTER) )
  this_clust <- subset(combi,cluster==CLUSTER)
  print(paste("total observations in cluster", CLUSTER," = ", nrow(this_clust) ))
  output <- dplyr::count(this_clust, Education, name=paste("Cluster", CLUSTER))
  print(output)
  
  write.table(output,
            "owner_edu_clusters.csv",
            sep=",",
            row.names=TRUE,
            append = TRUE)
}


```


#Pull out the centeroids
```{r centeroids}
p <- ggplot(df, aes(x, y)) +
    geom_point(aes(fill = Labs),alpha=0.8,shape=21,size=2, stroke=0.2) +
  labs(fill="Dog Diet")+
    xlab("UMAP X co-ordinates")+ylab("UMAP Y co-ordinates")+
    #geom_segment(aes(y=bottom,yend=top,x=-Inf,xend=-Inf))+
    #geom_segment(aes(y=-Inf,yend=-Inf,x=left,xend=right))+
    #scale_y_continuous(breaks=seq(bottom,top,2))+
    #scale_x_continuous(breaks=seq(left,r-ight,2))+
    geom_rangeframe() + theme_tufte(base_family = "Gill Sans") +
  geom_mark_circle(data=centeroids,
                     aes(x=mx, y=my, group=rownames(centeroids)), 
                   fill="white", 
                   alpha=0.7,
                   color="black",
                     con.type = "none",
                     label.fill = "white",
                     show.legend = FALSE,
                     inherit.aes = FALSE) +
  geom_text(data = centeroids, aes(mx, my, label = rownames(centeroids)), size = 10, color = "black")
    
    fname=paste0("umaps/","i", i,"min_dist",min_dist,
               "_spread" ,spread,"_nn" ,n_n, "_lcon" ,
               local_con,"_neg",neg,"_epochs",n_epochs, "_trees", n_trees,"_", scale,".png")
    print(p)
    #ggsave(fname)
```

#Calculate silhoutte and cluster metric
```{r silhoutte, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE,fig.showtext=TRUE}
set.seed(1234)

sil <- silhouette(db$cluster, dist(dplyr::select(combi,x,y)))
sil <- fviz_silhouette(sil)


sil_df <- sil$data
output <- data.frame(Cluster=as.numeric(), 
                     Score=as.numeric(),
                     mx=as.numeric(),
                     my=as.numeric()
                     )
mx <- 0
snx <- 0
for (cluster in unique(sil_df$cluster)){
  c <- as.character(cluster)
  Score <- mean(subset(sil_df,cluster==c)$sil_width)
  my <- Score
  #halfway long the cluster
  nx <- nrow(subset(sil_df,cluster==c))
  mx <- snx + nx/2
  snx <- snx + nx
  
  print(paste(cluster,round(Score,3), nx))
  local <- data.frame(
    Cluster=as.numeric(as.character(cluster)),
    Score=round(Score,3),
    mx=mx, my=my)
  output <- rbind(output,local)
}
head(output)
write.csv(output, "sils.csv")
output <- subset(output, Cluster >0)
p <-  sil + geom_mark_circle(data=output,
                  aes(x=mx, y=my, group=rownames(output)), 
                  fill="white", 
                  alpha=0.7,
                   color="black",
                     con.type = "none",
                     label.fill = "white",
                     show.legend = FALSE,
                     inherit.aes = FALSE) +
  geom_text(data = output, 
            aes(mx, my, label = round(Score,2)), 
            size = 3, color = "black")
print(p)
ggsave(paste0(figures,"fviz_silhouette.tiff"),width = 5, height = 4, dpi = 300, compression = "lzw")

```

#Davies-Bouldin
```{r cluster_metric, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE,fig.showtext=TRUE}
result <- index.DB(dplyr::select(combi,x,y), combi$cluster, centrotypes="centroids")

# print the Davies-Bouldin index
print(paste("Davies-Bouldin index is ", round(result$DB,4)) )
```

# UMAP by dog diet
```{r umap_dog,echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE,fig.showtext=TRUE}
Labs <- D_DietLabs[uKeeper] #remember we complete cased above!

# nice-ify labels
Labs <- stringr::str_replace(Labs,"Meat-based – conventional","Meat-based")
Labs <- stringr::str_replace(Labs,"Vegan \\(consuming no animal products\\)","Vegan")
Labs <- stringr::str_replace(Labs,"Meat-based – lab-grown","Lab-grown")
Labs <- stringr::str_replace(Labs,"Meat-based – raw","Raw food")
Labs <- stringr::str_replace(Labs,"Vegetarian \\(including eggs or milk, but not meat\\)","Vegetarian")

top=8
bottom=-6
left=-6
right=10
ggplot(df, aes(x, y)) +
  geom_point(aes(fill = Labs),
             alpha=0.8,shape=21,size=3, stroke=0.2) +
  labs(fill="Dog Diet")+
  xlab("UMAP X co-ordinates")+ylab("UMAP Y co-ordinates")+
  geom_segment(aes(y=bottom,yend=top,x=-Inf,xend=-Inf))+
  geom_segment(aes(y=-Inf,yend=-Inf,x=left,xend=right))+
  scale_fill_manual(values = rainbow(9)) +
  scale_y_continuous(breaks=seq(bottom,top,2))+
  scale_x_continuous(breaks=seq(left,right,2))+
  geom_rangeframe() + 
  theme_tufte(base_family = "Gill Sans", 
              base_size=18) +
  theme(legend.title = element_text(size = 18),
              legend.text = element_text(size = 12))+
  geom_rangeframe() + 
  geom_mark_circle(data=centeroids,
                     aes(x=mx, y=my, group=rownames(centeroids)), 
                   fill="white", 
                   alpha=0.3,
                   color="black",
                     con.type = "none",
                     label.fill = "white",
                     show.legend = FALSE,
                     inherit.aes = FALSE) +
  geom_text(data = centeroids, aes(mx, my, label = rownames(centeroids)), size = 8, color = "black")
ggsave(paste0(figures,"UMAP_Clear_Dog.TIFF"), compression="lzw")

```

# Initial Visualisation Excluding Health indicators with UMAP by extra metadata, for SI
```{r plot_other_UMAPS,echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE,fig.showtext=TRUE}
cropbraket <- function(mystring){
    newstring <- unlist(str_split(mystring, " \\("))[1]
  }

plotum <- function(labels,lcat,df){
  #number of legend rows
  lrow=2
  if (length(unique(labels))>3){lrow=3}
  if (length(unique(labels))>6){lrow=4}
  
  ldf <-cbind(df,labels)
  ldf <- ldf[complete.cases(ldf),]
  ldf <- ldf[order(ldf[,"labels"]),]
  labels<-ldf$labels

  #Pretty the legend key
  labels <- str_replace(labels,"Prefer not to answer","Not given")
  if (lcat=="D_Age") lcat="Dog Age"
  if (lcat=="C_Age") lcat="Owner Age"
  if (lcat=="Neuter_Status") lcat="Neuter Status"
  if (lcat=="D_Sex") lcat="Dog Sex"
  if (lcat=="C_Gender") lcat="Owner Gender"
  if (lcat=="C_Diet") {
      lcat="Owner Diet"
      # nice-ify labels
      labels <- stringr::str_replace(labels,
                                   "Vegetarian \\(consuming plants, eggs and milk\\)","Vegetarian")
      labels <- stringr::str_replace(labels,
                                   "Omnivore \\(i.e. eat meat\\)","Omnivore")
      labels <- stringr::str_replace(labels,
                                   "Omnivore reducing animal product consumption","Reduced Meat")
      labels <- stringr::str_replace(labels,
                                   "Vegan \\(consuming no animal products\\)" ,"Vegan")
      labels <- stringr::str_replace(labels,
                                   "Pescatarian \\(including fish but no other meats\\)","Pescatarian")
      order <- as.numeric(labels == "Vegan")
      ldf$order <- order
      ldf <- ldf[order(ldf[,"order"]),]
      
      #Try and get the colours to match the Dog equiv
      tempcols <- scales::hue_pal()(n=9)
      colours <- tempcols[4:9]
  }
  
  if (lcat=="D_Diet") return (NULL)
  
  if (lcat == "Animal_Career") {
    lcat = "Animal Career"
    labels <- str_replace(labels,
                            "None of the above","None")
    labels <- str_replace(labels,
                            "Pet industry worker","Pet industry")
    labels <- str_replace(labels,
                            "Animal trainer","Trainer")
    labels <- str_replace(labels,
                            "Veterinary technician or nurse","Vet Tech/Nurse")
    labels <- str_replace(labels,
                            "Animal breeder","Breeder")
  }
  
  if (lcat=="Size") labels <- sapply(labels,cropbraket)
  
  if (lcat=="Decision_Maker"){
      lcat="Decision Maker"
      labels <- str_replace(labels,
                            "Primary decision-maker","Primary")
      labels <- str_replace(labels,
                            "Play some lesser role","Some Role")
      labels <- str_replace(labels,
                            "Play no role","No Role")
  }
  
  if (lcat=="Location"){
      labels <- str_replace(labels,
                            "North America","N. America")
      labels <- str_replace(labels,
                            "Other European","Other Euro.")
      labels <- str_replace(labels,
                            "Australia/New Zealand/Oceania","Oceania")
  }

    if (lcat=="setting"){
      lcat="Setting"
      labels <- str_replace(labels,
                            "Equally urban and rural","Equal U & R")
  }
  
  if (lcat=="Education"){
      labels <- str_replace(labels,
                            "Did not complete high school","0_Basic")
      labels <- str_replace(labels,"High school or equivalent",
                            "1_HighSchool")
      labels <- str_replace(labels,
                            "College or University award lower than undergraduate degree",
                            "2_College")
      labels <- str_replace(labels,
                            "University undergraduate degree", "3_Grad")
      labels <- str_replace(labels,
                            "University postgraduate degree lower than doctoral degree e.g. masters", "4_PG")
      labels <- str_replace(labels,
                            "Doctoral degree", "5_PhD")
  }
  
    labels <- as.factor(str_trunc(labels,12,"right",""))
    if (lcat == "Dog Age") labels <- as.numeric(labels)
    #ldf <- dplyr::select(ldf,-labels)
  
  #Put the labels back in!
  ldf$labels<-labels

  #remove this version to remove confusion
  rm(labels)
  
  top=8
  bottom=-6
  left=-6
  right=10
 p <- ggplot()
 if (lcat == "Dog Diet") {
    p <- p + 
      geom_point(
        data = subset(ldf,!(labels %in% c("Vegan")) ),
               aes(x = x, y = y, fill = labels), 
               alpha = 0.8, shape = 21, size = 3, stroke = 0.2)+
      geom_point(data = subset(ldf, labels == "Vegan"),
              aes(x = x, y = y, fill = labels), alpha = 0.8, shape = 21, size = 3, stroke = 0.2)
 } else if (lcat == "Decision Maker") {
   p <- p + geom_point(
        data = subset(ldf,!(labels %in% c("No Role", "Some Role")) ),
               aes(x = x, y = y, fill = labels), 
               alpha = 0.8, shape = 21, size = 3, stroke = 0.2)+
      geom_point(data = subset(ldf, labels == "No Role"),
              aes(x = x, y = y, fill = labels), 
              alpha = 0.8, shape = 21, size = 3, stroke = 0.2)+
     geom_point(data = subset(ldf, labels == "Some Role"),
              aes(x = x, y = y, fill = labels), 
              alpha = 0.8, shape = 21, size = 3, stroke = 0.2)
 } else {
    p <- p + 
      geom_point(
        data = ldf,
               aes(x = x, y = y, fill = labels), 
               alpha = 0.8, shape = 21, size = 3, stroke = 0.2
        )
    }
 
  p <- p + labs(fill=lcat)+
    xlab("UMAP X co-ordinates")+ylab("UMAP Y co-ordinates")
    if (lcat=="Dog Age"){
      p <- p +scale_fill_gradientn(colors = rainbow(8)) 
    } else if (lcat == "Owner Diet"){
      p <- p +scale_fill_manual(values = colours,
                                   limits = unique(ldf$labels))
    }
  p <- p + geom_segment(aes(y=bottom,yend=top,x=-Inf,xend=-Inf))+
    geom_segment(aes(y=-Inf,yend=-Inf,x=left,xend=right))+
    scale_y_continuous(breaks=seq(bottom,top,2))+
    scale_x_continuous(breaks=seq(left,right,2))+
    geom_rangeframe() + 
    theme_tufte(base_family = "Gill Sans", base_size=18) +
    theme(legend.title = element_text(size = 18),
              legend.text = element_text(size = 12))+
    geom_mark_circle(data=centeroids,
                       aes(x=mx, y=my, 
                           group=rownames(centeroids)),
                       fill="white", 
                       alpha=0.3,
                       color="black",
                       con.type = "none",
                       label.fill = "white",
                       show.legend = FALSE,
                       inherit.aes = FALSE) +
    geom_text(data = centeroids, 
              aes(mx, my, 
                  label = rownames(centeroids)), 
              size = 8, color = "black")

  print(p)
  ggsave(paste0(figures,"UMAP_",lcat,".TIFF"), compression="lzw")
  return (ldf)
  }

for (label in names(Udata)){
  print(paste("starting",label))
  
  if (label=="Neuter_Status"){
    plotum(uml[["D_Neuter"]][uKeeper],label,df)
  } else{
      DF <- plotum(uml[[label]][uKeeper],label,df)
  }

  print(paste("completed",label) )
}
```


# Create ml data frame
```{r select_just_needed, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- dplyr::select(ml,c(unique.resp.number,
Location,
setting,
Education,
Animal_Career,
Income,
C_Age,
C_Gender,
C_Diet,
Size,
D_Age,
D_Gender,
D_Sex,
D_Neuter, 
Therapeutic_Food,
Visits,
Vet_Health,
Meds,
Health,
D_Diet,
Decision_Maker))

#check columns
head(ml)
```

# Categorise Fog Age into quintiles
```{r D_Age_quintiles, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <-subset(ml, D_Age != "NA")
ml$D_Age_quant <- cut(ml$D_Age , breaks=quantile(ml$D_Age, probs = seq(0, 1, 0.2)),na.rm = TRUE, include.lowest= TRUE)
dplyr::count(ml, D_Age_quant)
```


# Check Sex
```{r sex, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Sex <- as.factor(ml$D_Sex)
#Checks and Balances
dplyr::count(ml, D_Sex)
```

# Check Neuter
```{r neuter, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Neuter <- as.factor(ml$D_Neuter)
#Checks and Balances
dplyr::count(ml, D_Neuter)
```

# Check Location
```{r location, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Location)
```

# combine smaller groups into logical categories
```{r locations, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

for (i in 1:nrow(ml)){
  if (grepl('UK', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "UK"
  } else if (grepl('Other European', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "Other European"
  } else if (grepl('North America', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "North America"
  } else if (grepl('Australia/New Zealand/Oceania', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "Australia/New Zealand/Oceania"
  } else {
    ml$Location[i] <- "Other"
  }
}
ml$Location <- factor(ml$Location, levels = c("UK", "Other European", "North America", "Australia/New Zealand/Oceania", "Other"))
dplyr::count(ml, Location)
```

# create Location_Europe binary
```{r locationEur, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
# Create a binary Loaction_Europe variable
ml$Location_Europe <- ml$Location
ml$Location_Europe <- as.character(ml$Location_Europe)
for (i in 1:nrow(ml)){
  if (grepl('Other European', ml$Location_Europe[i], fixed = TRUE)){
    ml$Location_Europe[i] <- "Yes"
  } else {
    ml$Location_Europe[i] <- "No"
  }
}
dplyr::count(ml, Location_Europe)
```

# check setting
```{r settingcheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, setting)
```

# Not much point including data where setting is "other"
```{r groupraresettings, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, setting != "Other" | setting != NA)
ml$setting <- as.factor(ml$setting)
dplyr::count(ml, setting)
```

# Reorder SETTING and check
```{r reordersettings, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
# reorder categories so Urban is REF
ml$setting <- factor(ml$setting, levels=c('Urban', 'Rural', 'Equally urban and rural'))
dplyr::count(ml, setting)
```

# create binary for URBAN
```{r binary_urban, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
# Create a binary URBAN variable
ml$Urban <- ml$setting
ml$Urban <- as.character(ml$Urban)
for (i in 1:nrow(ml)){
  if (grepl('Urban', ml$Urban[i], fixed = TRUE)){
    ml$Urban[i] <- "Yes"
  } else {
    ml$Urban[i] <- "No"
  }
}
dplyr::count(ml, Urban)
```

# Check Education
```{r educheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Education)
```

# Simplify education and order categories
```{r simpedu, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Qualifications=c('0_Basic','1_HighSchool','2_College','3_Grad','4_PG', '5_PhD')
ml$Education <- str_replace(ml$Education,"Did not complete high school","0_Basic")
ml$Education <- str_replace(ml$Education,"High school or equivalent","1_HighSchool")
ml$Education <- str_replace(ml$Education,"College or University award lower than undergraduate degree", "2_College")
ml$Education <- str_replace(ml$Education,"University undergraduate degree", "3_Grad")
ml$Education <- str_replace(ml$Education,"University postgraduate degree lower than doctoral degree e.g. masters", "4_PG")
ml$Education <- str_replace(ml$Education,"Doctoral degree", "5_PhD")
ml$Education <- ordered(ml$Education)
dplyr::count(ml, Education)
```

#Simplify education further as small group sizes
```{r simpedu2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Qualifications_S=c('0_Basic_or_HighSchool','1_College','2_Grad','3_PG_or_PhD')
ml$Education_S <- ml$Education
ml$Education_S <- str_replace(ml$Education_S,"0_Basic","0_Basic_or_HighSchool")
ml$Education_S <- str_replace(ml$Education_S,"1_HighSchool","0_Basic_or_HighSchool")
ml$Education_S <- str_replace(ml$Education_S,"2_College", "1_College")
ml$Education_S <- str_replace(ml$Education_S,"3_Grad", "2_Grad")
ml$Education_S <- str_replace(ml$Education_S,"4_PG", "3_PG_or_PhD")
ml$Education_S <- str_replace(ml$Education_S,"5_PhD", "3_PG_or_PhD")
dplyr::count(ml, Education_S)
```

# Try non-ordered categories for Education_S
```{r non-ordered_group_education, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$Education_S2 <- factor(ml$Education_S, ordered=FALSE)
dplyr::count(ml, Education_S2)
```

# check Animal_Career variable
```{r animalcareercheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Animal_Career)
```

# create animal career variable where vet combined
```{r sort animalcareer, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$Animal_Career2 <- ml$Animal_Career
for (i in 1:nrow(ml)){
  if (grepl('Vet', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "Vet professional"
  } else if (grepl('Animal breeder', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "breeder/trainer"
  } else if (grepl('Animal trainer', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "breeder/trainer"
  } else if (grepl('Pet industry worker', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "Pet industry worker"
  } else {
    ml$Animal_Career2[i] <- "None of the above"
  }
}
#reorder categories so None of the above is REF
ml$Animal_Career2 <- factor(ml$Animal_Career2, levels=c('None of the above', 'Vet professional', 'breeder/trainer', 'Pet industry worker'))
dplyr::count(ml, Animal_Career2)
```

# create binary variable for animal career as ensures better balance
```{r binaryanimalcareer, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Animal_Career_BINARY <- ml$Animal_Career

for (i in 1:nrow(ml)){
  if (grepl('None of the above', ml$Animal_Career_BINARY[i], fixed = TRUE)){
    ml$Animal_Career_BINARY[i] <- "No"
  } else {
    ml$Animal_Career_BINARY[i] <- "Yes"
  }
}
dplyr::count(ml, Animal_Career_BINARY)
```

# Check Income
```{r checkincome, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Income) 
```

# Two choice with Income; treat as factor or drop "prefer not to say" and treat as ordinal.
```{r orderincome, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, Income != "Prefer not to answer" | Income != NA)
ml$Income <-ordered(ml$Income,levels=c("Low","Medium","High"))
dplyr::count(ml, Income)

```

# Try non-ordered categories for Income
```{r non-ordered_group_income, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$Income2 <- factor(ml$Income, ordered=FALSE)
dplyr::count(ml, Income2)
```

# Check C_Age
```{r check_C_ages, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, C_Age)
```

# combine smallest categories and reorder
```{r group_age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$C_Age <- str_replace(ml$C_Age,"18-19","<30")
ml$C_Age <- str_replace(ml$C_Age,"20–29","<30")
ml$C_Age <- str_replace(ml$C_Age,">70","60<")
ml$C_Age <- str_replace(ml$C_Age,"60–69","60<")
ml$C_Age<-ordered(ml$C_Age)
#Check right order
ml$C_Age[nrow(ml)]
dplyr::count(ml, C_Age)
```

# Try non-ordered categories for c_Age
```{r non-ordered_group_age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$C_Age2 <- factor(ml$C_Age, ordered=FALSE)
dplyr::count(ml, C_Age2)
```

# Check C_Gender
```{r check C-Gender, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, C_Gender) 
```

# Given small numbers in some categories, create male/female subset for C_Gender
```{r subset C_Gender, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <-subset(ml, C_Gender == "Male" | C_Gender =="Female")
ml$C_Gender <- as.factor(ml$C_Gender)
dplyr::count(ml, C_Gender) 
 
```

# Check C_Diet
```{r C_Diet, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, C_Diet) 
```

# Not much point including other category as small numbers
```{r remove_smallC_Diet_groups, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- subset(ml, C_Diet != "Other" | C_Diet != NA)
dplyr::count(ml, C_Diet) 
```

# create binary C_Vegan variable
```{r C_Vegan_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$C_Diet_Vegan <- ml$C_Diet
for (i in 1:nrow(ml)){
  if (grepl('Vegan', ml$C_Diet_Vegan[i], fixed = TRUE)){
    ml$C_Diet_Vegan[i] <- "Yes"
  } else {
    ml$C_Diet_Vegan[i] <- "No"
  }
}
dplyr::count(ml, C_Diet_Vegan)
```

# create binary C_Vegan_Veggie variable
```{r C_Vegan_Veggiebinary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$C_Diet_Vegan_Veggie <- ml$C_Diet
for (i in 1:nrow(ml)){
  if (grepl('Veg', ml$C_Diet_Vegan_Veggie[i], fixed = TRUE)){
    ml$C_Diet_Vegan_Veggie[i] <- "Yes"
  } else {
    ml$C_Diet_Vegan_Veggie[i] <- "No"
  }
}
dplyr::count(ml, C_Diet_Vegan_Veggie)
```

# reorder and rename size categories
```{r ordersizecategories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

for (i in 1:nrow(ml)){
  if (grepl('Toy', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Toy"
  } else if (grepl('Small', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Small"
  } else if (grepl('Medium', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Medium"
  } else if (grepl('Large', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Large"
  } else {
    ml$Size[i] <- "Giant"
  }
}
ml$Size <- ordered(ml$Size, levels=c("Toy","Small","Medium","Large","Giant"))
dplyr::count(ml, Size)
```

# Try non-ordered size categories with medium as reference
```{r non-orderedsizecategories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Size2 <- factor(ml$Size, levels=c("Medium","Toy","Small","Large","Giant"),ordered=FALSE)
dplyr::count(ml, Size2)
```

# create binary Giant breed variable
```{r create-Giant_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Giant=c('Yes','No')
ml$Size_Giant <- ml$Size2
ml$Size_Giant <- str_replace(ml$Size_Giant,"Giant","Yes")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Toy","No")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Medium","No")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Small","No")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Large","No")
dplyr::count(ml, Size_Giant)
```

# create theraeputic food category
```{r Therapeutic_food, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Therapeutic_Food <- as.character(ml$Therapeutic_Food)
ml <- subset(ml, Therapeutic_Food != "NA" | Therapeutic_Food != NA)
dplyr::count(ml, Therapeutic_Food)
```

# check visit data
```{r visit_data, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, Visits != "NA")
dplyr::count(ml, Visits)
```

# Not much point including data where clients are unsure of visit numbers.
```{r visitdata2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- subset(ml, Visits != "Unsure" | Visits != NA)
ml <- subset(ml, Visits != "NA" | Visits != NA)
dplyr::count(ml, Visits)
```

# rename and reorder Visits
```{r reordervisits, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Visits <- str_replace(ml$Visits,">3","3<")
ml$Visits <- str_replace(ml$Visits,"0","0")
ml$Visits <- str_replace(ml$Visits,"1","1")
ml$Visits <- str_replace(ml$Visits,"2","2")
ml$Visits <- str_replace(ml$Visits,"3","3")
dplyr::count(ml, ml$Visits)
```

# Simplify as Visits2 to ensure proportionality
```{r trysimplervisitcategories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Visits2 <- ml$Visits
ml$Visits2 <- str_replace(ml$Visits2,"3<","3")
ml$Visits2 <- str_replace(ml$Visits2,"3","3v")
ml$Visits2 <- str_replace(ml$Visits2,"0","0v")
ml$Visits2 <- str_replace(ml$Visits2,"1","1v")
ml$Visits2 <- str_replace(ml$Visits2,"2","2v")

dplyr::count(ml, ml$Visits2)
```

# check Meds
```{r Medscheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Meds <-as.factor(ml$Meds)
dplyr::count(ml, Meds)
```

# check Health category
```{r Healthcategorycheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, ml$Health)
```

# Tidy up Health category
## Not much point including data where clients unsure of pet health.
## Also need to reorder categories
```{r tidy_Health, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, Health != "Unsure" | Health != NA)
ml$Health <- as.factor(ml$Health)
ml$Health <-ordered(ml$Health,levels=c("Healthy","Generally healthy with minor or infrequent problems","Significant or frequent problems","Seriously ill"))
dplyr::count(ml, Health)
```

# create Health2 variable where serious and significant are combined
```{r Simplify_health_categories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Health_severity=c('0_None','1_Mild','2_Significant')
ml$Health2 <- ml$Health
ml$Health2 <- str_replace(ml$Health2,"Healthy","0_None")
ml$Health2 <- str_replace(ml$Health2,"Generally healthy with minor or infrequent problems","1_Mild")
ml$Health2 <- str_replace(ml$Health2,"Significant or frequent problems","2_Significant")
ml$Health2 <- str_replace(ml$Health2,"Seriously ill","2_Significant")
ml$Health2 <-ordered(ml$Health2,levels=c("0_None","1_Mild","2_Significant"))
dplyr::count(ml, ml$Health2)
```

# create binary Health variable
```{r create_health_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
bhealth <- rep(0,nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('healthy', ml$Health[i], ignore.case = TRUE)){
      # 0 is good health 
      bhealth[i] <- 0
       }
    else {
       bhealth[i] <- 1
    }
}
#Keep as a variable not factor!
ml$Health_Binary <- as.numeric(bhealth)
rm(bhealth)
#Checks and Balances, hey it's VERY unbalanced
dplyr::count(ml, Health_Binary)

```

# create binary Any health problem variable
```{r create_any health_problem_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
bhealth2 <- rep(0,nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('0_None', ml$Health2[i], ignore.case = TRUE)){
      # 0 is good health 
      bhealth2[i] <- 0
       }
    else {
       bhealth2[i] <- 1
    }
}
#Keep as a variable not factor!
ml$Any_Health_Problem <- as.numeric(bhealth2)
rm(bhealth2)
#Checks and Balances, better balanced this time.
dplyr::count(ml, Any_Health_Problem)

```

# check D_Diet
```{r checkD_Diet, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, D_Diet)
```

# Not much point including those that are unsure.
```{r tidy__Diet, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- subset(ml, D_Diet != "Unsure" | D_Diet != NA)
dplyr::count(ml, D_Diet) 
```

# Not much point including diets where there are small numbers
```{r tidy_D_Diet2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, D_Diet != "Mixture" | D_Diet != NA)
ml <- subset(ml, D_Diet != "Insect-based" | D_Diet != NA)
ml <- subset(ml, D_Diet != "Meat-based – lab-grown" | D_Diet != NA)
dplyr::count(ml, D_Diet) 
```

# create binary D_Vegan variable
```{r create_D_Vegan binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Diet_Vegan <- ml$D_Diet
for (i in 1:nrow(ml)){
  if (grepl('Vegan', ml$D_Diet_Vegan[i], fixed = TRUE)){
    ml$D_Diet_Vegan[i] <- "Yes"
  } else {
    ml$D_Diet_Vegan[i] <- "No"
  }
}
dplyr::count(ml, D_Diet_Vegan)
```

# create binary D_Raw variable
```{r create_D_Raw_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Diet_Raw <- ml$D_Diet
for (i in 1:nrow(ml)){
  if (grepl('raw', ml$D_Diet_Raw[i], fixed = TRUE)){
    ml$D_Diet_Raw[i] <- "Yes"
  } else {
    ml$D_Diet_Raw[i] <- "No"
  }
}
dplyr::count(ml, D_Diet_Raw)
```

# create binary D_Vegan_Veggie variable
```{r create_D_Vegan-Veggie_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Diet_Vegan_Veggie <- ml$D_Diet
for (i in 1:nrow(ml)){
  if (grepl('Veg', ml$D_Diet_Vegan_Veggie[i], fixed = TRUE)){
    ml$D_Diet_Vegan_Veggie[i] <- "Yes"
  } else {
    ml$D_Diet_Vegan_Veggie[i] <- "No"
  }
}
dplyr::count(ml, D_Diet_Vegan_Veggie)
```

# check Decision_maker
```{r check_Decision_Maker, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Decision_Maker)
```

# Better only to include data from clients that were primary decision makers as likely to be more accurate.
```{r tidy_Decision_Maker, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, Decision_Maker == "Primary decision-maker")
dplyr::count(ml, Decision_Maker)
```

# Let's just keep those we will need
```{r final_column_check, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml,select = c("Location", "setting","Urban", "Education", "Education_S","Education_S2", "Animal_Career",
                           "Animal_Career2","Animal_Career_BINARY", "Income","Income2","C_Age", "C_Age2", "C_Gender",
                           "C_Diet", "C_Diet_Vegan", "C_Diet_Vegan_Veggie", "D_Age", "D_Age_quant",
                           "Size", "Size2", "Size_Giant", "D_Sex", "D_Neuter", "Therapeutic_Food","Meds", "D_Diet",
                           "D_Diet_Vegan", "D_Diet_Vegan_Veggie", "D_Diet_Raw", "Visits", "Visits2", "Health",
                           "Health2","Health_Binary", "Any_Health_Problem", "Location_Europe"))

#check columns
head(ml)
```

# summary data
```{r summarise, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
describe(ml)
```

# check C_Diet category numbers
```{r C_Diet_check, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml,C_Diet)
```

# check D_Sex by D_Neuter
```{r D_2sex_neuter_check, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
## two way cross tabs (xtabs) and flatten the table
ftable(xtabs(~ D_Sex + D_Neuter, data = ml))
```

# Finally, create train and test datasets for use in logistic regression analyses
```{r train_test}
index <- sample(nrow(ml),nrow(ml)*0.75)
# set seed to make partition
set.seed(123)
ml_train = ml[index,]
ml_test = ml[-index,]
```

# Citations for packages
```{r citations, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
citation("stringr")
packageVersion("stringr")
citation("dplyr")
packageVersion("dplyr")
citation("DescTools")
packageVersion("DescTools")
citation("car")
packageVersion("car")
citation("foreign")
packageVersion("foreign")
citation("ggplot2")
packageVersion("ggplot2")
citation("Hmisc")
packageVersion("Hmisc")
citation("reshape2")
packageVersion("reshape2")
citation("tidyverse")
packageVersion("tidyverse")
citation("broom")
packageVersion("broom")
citation("aod")
packageVersion("aod")
citation("glmtoolbox")
packageVersion("glmtoolbox")
citation("fmsb")
packageVersion("fmsb")
citation("uwot")
packageVersion("uwot")
citation("ggthemes")
packageVersion("ggthemes")
citation("ggforce")
packageVersion("ggforce")
citation("readxl")
packageVersion("readxl")
citation("ROCR")
packageVersion("ROCR")
citation("pROC")
packageVersion("pROC")
citation("caret")
packageVersion("caret")
citation("xgboost")
packageVersion("xgboost")
citation("smotefamily")
packageVersion("smotefamily")
citation("rsample")
packageVersion("rsample")
citation("mltools")
packageVersion("mltools")
citation("data.table")
packageVersion("data.table")
citation("heatmaply")
packageVersion("heatmaply")
citation("rstatix")
packageVersion("rstatix")
citation("dbscan")
packageVersion("dbscan")
citation("corrplot")
packageVersion("corrplot")
citation("showtext")
packageVersion("showtext")
citation("cluster")
packageVersion("cluster")
citation("factoextra")
packageVersion("factoextra")
citation("clusterSim")
packageVersion("clusterSim")
citation("RColorBrewer")
packageVersion("RColorBrewer")
citation("IHW")
packageVersion("IHW")
```

```{r savedata, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
if (SAVE){
  save(ml, file = "ml_read_data_101.RData")
}
```


