---
title: "Read Data 102 - include all owners not just decision makers"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

Data from here:
https://osf.io/8kzfv?view_only=d05ee82e58dc4c0f84448785ab839bb3
Original methods and full paper here:
https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0265662

# Load required packages
```{r loadpkgs, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE, results='hide'}
library(stringr)
library(dplyr)
library(DescTools)
library(car)
require(foreign)
require(ggplot2)
require(Hmisc)
require(reshape2)
library(tidyverse)
library(broom)
library(aod)
library(glmtoolbox)
library(fmsb)
library(umap)
library(ggthemes)
library(readxl) #leaves characters as characters
library(ROCR) # Use ROCR package to plot ROC curve & AUC
library(pROC)
library(caret)
library(xgboost)
library(smotefamily) # one method of ML data augmentation
library(rsample)
library(mltools) # Wrapper allows UMAP calculations
library(data.table)
library(heatmaply)
library(rstatix) # to calculate Kendall correlation with associated p-values.
library(IHW) # Bioconductor package to calculate BH threshold
library(dbscan)
library(corrplot)
library(showtext)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
library(splines)
library(splines)
library(cluster)
library(factoextra)
library(clusterSim)
```

# Open Excel file
```{r read_data_1, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
#Save the data so can run subsequent analysis ad lib
SAVE=TRUE

#quicker version readxl
ml <- read_excel('Canine health results.xlsx', sheet = "All")
cols <- c("Col_a")
for (col in 2:ncol(ml)){
  if (col <= 26){
    cols <- cbind(cols,paste0("Col_",letters[col]))
  } else if (col <= 52) {
    cols <- cbind(cols,paste0("Col_a",letters[col-26]))
  } else {
    cols <- cbind(cols,paste0("Col_b",letters[col-52]))
  }
}
names(ml) <- cols

```

# Rename variables
```{r RENAME_MATCH_2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- dplyr::rename(ml, unique.resp.number = Col_a)
ml <- dplyr::rename(ml, Location = Col_c)
ml <- dplyr::rename(ml, setting = Col_e)
ml <- dplyr::rename(ml, Education = Col_g)
ml <- dplyr::rename(ml, Animal_Career = Col_h)
ml <- dplyr::rename(ml, Income = Col_i)
ml <- dplyr::rename(ml, C_Age = Col_j)
ml <- dplyr::rename(ml, C_Gender = Col_k)
ml <- dplyr::rename(ml, C_Diet = Col_l)
ml <- dplyr::rename(ml, Size = Col_o)
ml <- dplyr::rename(ml, D_Age = Col_p)
ml <- dplyr::rename(ml, D_Gender = Col_r)
ml <- dplyr::rename(ml, Therapeutic_Food = Col_s)
ml <- dplyr::rename(ml, Visits = Col_t)
ml <- dplyr::rename(ml, Vet_Health = Col_u)
ml <- dplyr::rename(ml, Meds = Col_ag)
ml <- dplyr::rename(ml, Health = Col_ah)
ml <- dplyr::rename(ml, D_Diet = Col_ai)
ml <- dplyr::rename(ml, Decision_Maker = Col_at)
head(ml)
```



# Start by creating separate Sex and neuter status variables
```{r split_gender_Neuter, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Sex <- rep("Male",nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('Male', ml$D_Gender[i], fixed = TRUE)){
       ml$D_Sex[i] <- "Male"
       } else{
       ml$D_Sex[i] <- "Female"
       }
}

#Checks and Balances
dplyr::count(ml, D_Sex)

ml$D_Neuter <- rep("BLANK",nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('intact', ml$D_Gender[i], fixed = TRUE)){
       ml$D_Neuter[i] <- "Intact"
       }
    else {
       ml$D_Neuter[i] <- "Neutered"
    }
}
dplyr::count(ml, D_Neuter)
```

# check D_Age
```{r check_D_Age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, D_Age)
```

# Dog Age remove <1 years and unsure, then change to numeric
```{r tidy_D_Age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <-subset(ml, D_Age != "<1" & D_Age != "Unsure")
ml$D_Age<-as.numeric(ml$D_Age)
dplyr::count(ml, D_Age)
```

# Create ml data frame
```{r select_just_needed, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- dplyr::select(ml,c(unique.resp.number,
Location,
setting,
Education,
Animal_Career,
Income,
C_Age,
C_Gender,
C_Diet,
Size,
D_Age,
D_Gender,
D_Sex,
D_Neuter, 
Therapeutic_Food,
Visits,
Vet_Health,
Meds,
Health,
D_Diet,
Decision_Maker))

#check columns
head(ml)
```

# Categorise Dog Age into quintiles
```{r D_Age_quintiles, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <-subset(ml, D_Age != "NA")
ml$D_Age_quant <- cut(ml$D_Age , breaks=quantile(ml$D_Age, probs = seq(0, 1, 0.2)),na.rm = TRUE, include.lowest = TRUE)
dplyr::count(ml, D_Age_quant)
```



# Check Sex
```{r sex, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Sex <- as.factor(ml$D_Sex)
#Checks and Balances
dplyr::count(ml, D_Sex)
```

# Check Neuter
```{r neuter, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Neuter <- as.factor(ml$D_Neuter)
#Checks and Balances
dplyr::count(ml, D_Neuter)
```

# Check Location
```{r location, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Location)
```

# combine smaller groups into logical categories
```{r locations, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

for (i in 1:nrow(ml)){
  if (grepl('UK', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "UK"
  } else if (grepl('Other European', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "Other European"
  } else if (grepl('North America', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "North America"
  } else if (grepl('Australia/New Zealand/Oceania', ml$Location[i], fixed = TRUE)){
    ml$Location[i] <- "Australia/New Zealand/Oceania"
  } else {
    ml$Location[i] <- "Other"
  }
}
ml$Location <- factor(ml$Location, levels = c("UK", "Other European", "North America", "Australia/New Zealand/Oceania", "Other"))
dplyr::count(ml, Location)
```

# check setting
```{r settingcheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, setting)
```

# Not much point including data where setting is "other"
```{r groupraresettings, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, setting != "Other" | setting != NA)
ml$setting <- as.factor(ml$setting)
dplyr::count(ml, setting)
```

# Reorder SETTING and check
```{r reordersettings, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
# reorder categories so Urban is REF
ml$setting <- factor(ml$setting, levels=c('Urban', 'Rural', 'Equally urban and rural'))
dplyr::count(ml, setting)
```

# create binary for URBAN
```{r binary_urban, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
# Create a binary URBAN variable
ml$Urban <- ml$setting
ml$Urban <- as.character(ml$Urban)
for (i in 1:nrow(ml)){
  if (grepl('Urban', ml$Urban[i], fixed = TRUE)){
    ml$Urban[i] <- "Yes"
  } else {
    ml$Urban[i] <- "No"
  }
}
dplyr::count(ml, Urban)
```

# Check Education
```{r educheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Education)
```

# Simplify education and order categories
```{r simpedu, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Qualifications=c('0_Basic','1_HighSchool','2_College','3_Grad','4_PG', '5_PhD')
ml$Education <- str_replace(ml$Education,"Did not complete high school","0_Basic")
ml$Education <- str_replace(ml$Education,"High school or equivalent","1_HighSchool")
ml$Education <- str_replace(ml$Education,"College or University award lower than undergraduate degree", "2_College")
ml$Education <- str_replace(ml$Education,"University undergraduate degree", "3_Grad")
ml$Education <- str_replace(ml$Education,"University postgraduate degree lower than doctoral degree e.g. masters", "4_PG")
ml$Education <- str_replace(ml$Education,"Doctoral degree", "5_PhD")
ml$Education <- ordered(ml$Education)
dplyr::count(ml, Education)
```

#Simplify education further as small group sizes
```{r simpedu2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Qualifications_S=c('0_Basic_or_HighSchool','1_College','2_Grad','3_PG_or_PhD')
ml$Education_S <- ml$Education
ml$Education_S <- str_replace(ml$Education_S,"0_Basic","0_Basic_or_HighSchool")
ml$Education_S <- str_replace(ml$Education_S,"1_HighSchool","0_Basic_or_HighSchool")
ml$Education_S <- str_replace(ml$Education_S,"2_College", "1_College")
ml$Education_S <- str_replace(ml$Education_S,"3_Grad", "2_Grad")
ml$Education_S <- str_replace(ml$Education_S,"4_PG", "3_PG_or_PhD")
ml$Education_S <- str_replace(ml$Education_S,"5_PhD", "3_PG_or_PhD")
dplyr::count(ml, Education_S)
```


# Try non-ordered categories for Education_S
```{r non-ordered_group_education, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$Education_S2 <- factor(ml$Education_S, ordered=FALSE)
dplyr::count(ml, Education_S2)
```


# check Animal_Career variable
```{r animalcareercheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Animal_Career)
```

# create animal career variable where vet combined
```{r sort animalcareer, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$Animal_Career2 <- ml$Animal_Career
for (i in 1:nrow(ml)){
  if (grepl('Vet', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "Vet professional"
  } else if (grepl('Animal breeder', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "breeder/trainer"
  } else if (grepl('Animal trainer', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "breeder/trainer"
  } else if (grepl('Pet industry worker', ml$Animal_Career2[i], fixed = TRUE)){
    ml$Animal_Career2[i] <- "Pet industry worker"
  } else {
    ml$Animal_Career2[i] <- "None of the above"
  }
}
#reorder categories so None of the above is REF
ml$Animal_Career2 <- factor(ml$Animal_Career2, levels=c('None of the above', 'Vet professional', 'breeder/trainer', 'Pet industry worker'))
dplyr::count(ml, Animal_Career2)
```


# create binary variable for animal career as ensures better balance
```{r binaryanimalcareer, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Animal_Career_BINARY <- ml$Animal_Career

for (i in 1:nrow(ml)){
  if (grepl('None of the above', ml$Animal_Career_BINARY[i], fixed = TRUE)){
    ml$Animal_Career_BINARY[i] <- "No"
  } else {
    ml$Animal_Career_BINARY[i] <- "Yes"
  }
}
dplyr::count(ml, Animal_Career_BINARY)
```

# Check Income
```{r checkincome, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Income) 
```

# Two choice with Income; treat as factor or drop "prefer not to say" and treat as ordinal.
```{r orderincome, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, Income != "Prefer not to answer" | Income != NA)
ml$Income <-ordered(ml$Income,levels=c("Low","Medium","High"))
dplyr::count(ml, Income)

```


# Try non-ordered categories for Income
```{r non-ordered_group_income, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$Income2 <- factor(ml$Income, ordered=FALSE)
dplyr::count(ml, Income2)
```


# Check C_Age
```{r check_C_ages, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, C_Age)
```

# combine smallest categories and reorder
```{r group_age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$C_Age <- str_replace(ml$C_Age,"18-19","<30")
ml$C_Age <- str_replace(ml$C_Age,"20–29","<30")
ml$C_Age <- str_replace(ml$C_Age,">70","60<")
ml$C_Age <- str_replace(ml$C_Age,"60–69","60<")
ml$C_Age<-ordered(ml$C_Age)
#Check right order
ml$C_Age[nrow(ml)]
dplyr::count(ml, C_Age)
```


# Try non-ordered categories for c_Age
```{r non-ordered_group_age, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$C_Age2 <- factor(ml$C_Age, ordered=FALSE)
dplyr::count(ml, C_Age2)
```


# Check C_Gender
```{r check C-Gender, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, C_Gender) 
```

# Given small numbers in some categories, create male/female subset for C_Gender
```{r subset C_Gender, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <-subset(ml, C_Gender == "Male" | C_Gender =="Female")
ml$C_Gender <- as.factor(ml$C_Gender)
dplyr::count(ml, C_Gender) 
 
```

# Check C_Diet
```{r C_Diet, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, C_Diet) 
```

# Not much point including other category as small numbers
```{r remove_smallC_Diet_groups, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- subset(ml, C_Diet != "Other" | C_Diet != NA)
dplyr::count(ml, C_Diet) 
```

# create binary C_Vegan variable
```{r C_Vegan_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$C_Diet_Vegan <- ml$C_Diet
for (i in 1:nrow(ml)){
  if (grepl('Vegan', ml$C_Diet_Vegan[i], fixed = TRUE)){
    ml$C_Diet_Vegan[i] <- "Yes"
  } else {
    ml$C_Diet_Vegan[i] <- "No"
  }
}
dplyr::count(ml, C_Diet_Vegan)
```

# create binary C_Vegan_Veggie variable
```{r C_Vegan_Veggiebinary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml$C_Diet_Vegan_Veggie <- ml$C_Diet
for (i in 1:nrow(ml)){
  if (grepl('Veg', ml$C_Diet_Vegan_Veggie[i], fixed = TRUE)){
    ml$C_Diet_Vegan_Veggie[i] <- "Yes"
  } else {
    ml$C_Diet_Vegan_Veggie[i] <- "No"
  }
}
dplyr::count(ml, C_Diet_Vegan_Veggie)
```

# reorder and rename size categories
```{r ordersizecategories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

for (i in 1:nrow(ml)){
  if (grepl('Toy', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Toy"
  } else if (grepl('Small', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Small"
  } else if (grepl('Medium', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Medium"
  } else if (grepl('Large', ml$Size[i], fixed = TRUE)){
    ml$Size[i] <- "Large"
  } else {
    ml$Size[i] <- "Giant"
  }
}
ml$Size <- ordered(ml$Size, levels=c("Toy","Small","Medium","Large","Giant"))
dplyr::count(ml, Size)
```


# Try non-ordered size categories with medium as reference
```{r non-orderedsizecategories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Size2 <- factor(ml$Size, levels=c("Medium","Toy","Small","Large","Giant"),ordered=FALSE)
dplyr::count(ml, Size2)
```


# create binary Giant breed variable
```{r create-Giant_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Giant=c('Yes','No')
ml$Size_Giant <- ml$Size2
ml$Size_Giant <- str_replace(ml$Size_Giant,"Giant","Yes")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Toy","No")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Medium","No")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Small","No")
ml$Size_Giant <- str_replace(ml$Size_Giant,"Large","No")
dplyr::count(ml, Size_Giant)
```


# create theraeputic food category
```{r Therapeutic_food, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Therapeutic_Food <- as.character(ml$Therapeutic_Food)
ml <- subset(ml, Therapeutic_Food != "NA" | Therapeutic_Food != NA)
dplyr::count(ml, Therapeutic_Food)
```

# check visit data
```{r visit_data, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, Visits != "NA")
dplyr::count(ml, Visits)
```

# Not much point including data where clients are unsure of visit numbers.
```{r visitdata2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- subset(ml, Visits != "Unsure" | Visits != NA)
ml <- subset(ml, Visits != "NA" | Visits != NA)
dplyr::count(ml, Visits)
```

# rename and reorder Visits
```{r reordervisits, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Visits <- str_replace(ml$Visits,">3","3<")
ml$Visits <- str_replace(ml$Visits,"0","0")
ml$Visits <- str_replace(ml$Visits,"1","1")
ml$Visits <- str_replace(ml$Visits,"2","2")
ml$Visits <- str_replace(ml$Visits,"3","3")
dplyr::count(ml, ml$Visits)
```


# Simplify as Visits2 to ensure proportionality
```{r trysimplervisitcategories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Visits2 <- ml$Visits
ml$Visits2 <- str_replace(ml$Visits2,"3<","3")
ml$Visits2 <- str_replace(ml$Visits2,"3","3v")
ml$Visits2 <- str_replace(ml$Visits2,"0","0v")
ml$Visits2 <- str_replace(ml$Visits2,"1","1v")
ml$Visits2 <- str_replace(ml$Visits2,"2","2v")

dplyr::count(ml, ml$Visits2)
```

# check Meds
```{r Medscheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Meds <-as.factor(ml$Meds)
dplyr::count(ml, Meds)
```

# check Health category
```{r Healthcategorycheck, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, ml$Health)
```


# Tidy up Health category
## Not much point including data where clients unsure of pet health.
## Also need to reorder categories
```{r tidy_Health, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, Health != "Unsure" | Health != NA)
ml$Health <- as.factor(ml$Health)
ml$Health <-ordered(ml$Health,levels=c("Healthy","Generally healthy with minor or infrequent problems","Significant or frequent problems","Seriously ill"))
dplyr::count(ml, Health)
```

# create Health2 variable where serious and significant are combined
```{r Simplify_health_categories, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
Health_severity=c('0_None','1_Mild','2_Significant')
ml$Health2 <- ml$Health
ml$Health2 <- str_replace(ml$Health2,"Healthy","0_None")
ml$Health2 <- str_replace(ml$Health2,"Generally healthy with minor or infrequent problems","1_Mild")
ml$Health2 <- str_replace(ml$Health2,"Significant or frequent problems","2_Significant")
ml$Health2 <- str_replace(ml$Health2,"Seriously ill","2_Significant")
ml$Health2 <-ordered(ml$Health2,levels=c("0_None","1_Mild","2_Significant"))
dplyr::count(ml, ml$Health2)
```


# create binary Health variable
```{r create_health_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
bhealth <- rep(0,nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('healthy', ml$Health[i], ignore.case = TRUE)){
      # 0 is good health 
      bhealth[i] <- 0
       }
    else {
       bhealth[i] <- 1
    }
}
#Keep as a variable not factor!
ml$Health_Binary <- as.numeric(bhealth)
rm(bhealth)
#Checks and Balances, hey it's VERY unbalanced
dplyr::count(ml, Health_Binary)

```


# create binary Any health problem variable
```{r create_any health_problem_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
bhealth2 <- rep(0,nrow(ml))
for (i in 1:nrow(ml)){
    if (grepl('0_None', ml$Health2[i], ignore.case = TRUE)){
      # 0 is good health 
      bhealth2[i] <- 0
       }
    else {
       bhealth2[i] <- 1
    }
}
#Keep as a variable not factor!
ml$Any_Health_Problem <- as.numeric(bhealth2)
rm(bhealth2)
#Checks and Balances, better balanced this time.
dplyr::count(ml, Any_Health_Problem)

```


# check D_Diet
```{r checkD_Diet, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, D_Diet)
```


# Not much point including those that are unsure.
```{r tidy__Diet, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}

ml <- subset(ml, D_Diet != "Unsure" | D_Diet != NA)
dplyr::count(ml, D_Diet) 
```


# Not much point including diets where there are small numbers
```{r tidy_D_Diet2, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml, D_Diet != "Mixture" | D_Diet != NA)
ml <- subset(ml, D_Diet != "Insect-based" | D_Diet != NA)
ml <- subset(ml, D_Diet != "Meat-based – lab-grown" | D_Diet != NA)
dplyr::count(ml, D_Diet) 
```


# create binary D_Vegan variable
```{r create_D_Vegan binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Diet_Vegan <- ml$D_Diet
for (i in 1:nrow(ml)){
  if (grepl('Vegan', ml$D_Diet_Vegan[i], fixed = TRUE)){
    ml$D_Diet_Vegan[i] <- "Yes"
  } else {
    ml$D_Diet_Vegan[i] <- "No"
  }
}
dplyr::count(ml, D_Diet_Vegan)
```


# create binary D_Raw variable
```{r create_D_Raw_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Diet_Raw <- ml$D_Diet
for (i in 1:nrow(ml)){
  if (grepl('raw', ml$D_Diet_Raw[i], fixed = TRUE)){
    ml$D_Diet_Raw[i] <- "Yes"
  } else {
    ml$D_Diet_Raw[i] <- "No"
  }
}
dplyr::count(ml, D_Diet_Raw)
```


# create binary D_Vegan_Veggie variable
```{r create_D_Vegan-Veggie_binary, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$D_Diet_Vegan_Veggie <- ml$D_Diet
for (i in 1:nrow(ml)){
  if (grepl('Veg', ml$D_Diet_Vegan_Veggie[i], fixed = TRUE)){
    ml$D_Diet_Vegan_Veggie[i] <- "Yes"
  } else {
    ml$D_Diet_Vegan_Veggie[i] <- "No"
  }
}
dplyr::count(ml, D_Diet_Vegan_Veggie)
```


# check Decision_maker
```{r check_Decision_Maker, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml, Decision_Maker)
```

# create binary for primary decision maker
```{r create-decision_maker, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml$Primary_Decision_Maker <- ml$Decision_Maker
for (i in 1:nrow(ml)){
  if (grepl('Primary decision-maker',ml$Primary_Decision_Maker[i], fixed = TRUE)){
    ml$Primary_Decision_Maker[i] <- "Yes"
  } else {
   ml$Primary_Decision_Maker[i] <- "No"
  }
}
dplyr::count(ml, ml$Primary_Decision_Maker)
```


# Let's just keep those we will need
```{r final_column_check, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
ml <- subset(ml,select = c("Location", "setting","Urban", "Education", "Education_S","Education_S2", "Animal_Career",
                           "Animal_Career2","Animal_Career_BINARY", "Income","Income2","C_Age", "C_Age2", "C_Gender",
                           "C_Diet", "C_Diet_Vegan", "C_Diet_Vegan_Veggie", "D_Age", "D_Age_quant",
                           "Size", "Size2", "Size_Giant", "D_Sex", "D_Neuter", "Therapeutic_Food","Meds", "D_Diet",
                           "D_Diet_Vegan", "D_Diet_Vegan_Veggie", "D_Diet_Raw", "Visits", "Visits2", "Health",
                           "Health2","Health_Binary", "Any_Health_Problem","Primary_Decision_Maker"))

#check columns
head(ml)
```





# summary data
```{r summarise, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
describe(ml)
```


# check C_Diet category numbers
```{r C_Diet_check, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
dplyr::count(ml,C_Diet)
```


# check D_Sex by D_Neuter
```{r D_2sex_neuter_check, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
## two way cross tabs (xtabs) and flatten the table
ftable(xtabs(~ D_Sex + D_Neuter, data = ml))
```


# Citations for packages
```{r citations, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
citation("stringr")
packageVersion("stringr")
citation("dplyr")
packageVersion("dplyr")
citation("DescTools")
packageVersion("DescTools")
citation("car")
packageVersion("car")
citation("foreign")
packageVersion("foreign")
citation("ggplot2")
packageVersion("ggplot2")
citation("Hmisc")
packageVersion("Hmisc")
citation("reshape2")
packageVersion("reshape2")
citation("tidyverse")
packageVersion("tidyverse")
citation("broom")
packageVersion("broom")
citation("aod")
packageVersion("aod")
citation("glmtoolbox")
packageVersion("glmtoolbox")
citation("fmsb")
packageVersion("fmsb")
citation("uwot")
packageVersion("uwot")
citation("ggthemes")
packageVersion("ggthemes")
citation("ggforce")
packageVersion("ggforce")
citation("readxl")
packageVersion("readxl")
citation("ROCR")
packageVersion("ROCR")
citation("pROC")
packageVersion("pROC")
citation("caret")
packageVersion("caret")
citation("xgboost")
packageVersion("xgboost")
citation("smotefamily")
packageVersion("smotefamily")
citation("rsample")
packageVersion("rsample")
citation("mltools")
packageVersion("mltools")
citation("data.table")
packageVersion("data.table")
citation("heatmaply")
packageVersion("heatmaply")
citation("rstatix")
packageVersion("rstatix")
citation("dbscan")
packageVersion("dbscan")
citation("corrplot")
packageVersion("corrplot")
citation("showtext")
packageVersion("showtext")
citation("cluster")
packageVersion("cluster")
citation("factoextra")
packageVersion("factoextra")
citation("clusterSim")
packageVersion("clusterSim")
citation("RColorBrewer")
packageVersion("RColorBrewer")
citation("IHW")
packageVersion("IHW")
```

```{r savedata, echo=TRUE, include=TRUE, warning=FALSE, errors=FALSE, message=FALSE}
if (SAVE){
  save(ml, file = "ml_read_data_102.RData")
}
```


